Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-04-30 GWLDCII.tex
minor progress on paper 2

@@

We summarize here only the subset of positroid theory that we require in this paper; the interested reader is referred to \cite{Postnikov} for more details. Postnikov shows in \cite{Postnikov} that the positroids of $\Gr(k,n)$ are indexed by many different collections of objects, each with their own advantages and disadvantages. The two most suited to our purposes are Grassmann necklaces and Le diagrams, which we introduce below. In order to do so, we first need some preliminary definitions.

@@

In \cite{casestudy}, Agarwala and Fryer apply this process in the smallest non-trivial case: admissible Wilson loop diagrams with two propagators on six vertices. Even in this simple case, we see that the mapping from admissible WLD to positroid cells is neither one-to-one nor onto, and the relationship between the original Wilson loop diagram and the resulting positroid is obscured to the point of invisibility.

@@

Given a propagator $p = (i, j)$ oriented $(p, <_i)$,  we can construct two obvious subdiagrams of $W$ by removing all of the propagators on one side of $p$ and leaving the vertices intact, i.e. the subdiagrams
\[  (\cP_{in}\cup\{p\}, [n])  \quad \textrm{and} \quad  (\cP_{out}\cup \{p\}, [n]).\]

%where the set of non-supportive vertices of  $W|_{\cP_{in} \cup p}$ is $V_{out}$ and  the set of non-supportive vertices of  $W|_{\cP_{out} \cup p}$ is $V_{in}$. That is, these restrictions contain the propagators on the inside and outside of $(p, <_i)$ respectively, as well as $p$ itself. 
%\footnote{If we denote $V \subset [n]$ and $Q = \Prop(V^c)$, then we can write \bas M(W|_{Q^c}) = M(W)/F(Q) \oplus M_{\emptyset, n-|i-j+1|} \eas where  $M_{\emptyset, k}$ is the matroid of rank $0$ on a set of $k$. That is $M_{\emptyset, n-|i-j+1|} =M((\emptyset, n-|i-j+1|))$. }

%We say that $p$ \textit{bounds the regions} $[i,j+1]$ and $[j,i+1]$ which we call outside and inside respectively. The ordering of the start and end points allow us to identify one or the other region unambiguously.

%With these observations in mind we can return to the proof.

We will inductively construct a sequence of pairs of propagators $(p_r,q_r)$ satisfying: $\ell(p_r) = 2$, and $p_r$ either forms part of configuration \ref{item big and 2} or \ref{item pair of 2s} from the statement,
  %a configuration \eqref{eq:2 prop config} in $D$
or there is a propagator $q_r$ satisfying
\begin{itemize}
\item $\ell(q_r) \geq 4$.
\item $\{p_1, \ldots, p_r\}$ and $\{q_1, \ldots, q_{r-1}\}$ are all on the same side of $q_r$.  
\end{itemize}
We can then choose the orientation of $q_r$ so that the previous $p_i$ and $q_i$ are all on the outside of $q_r$, and 
By the finiteness of $W$ this must eventually terminate in one of the desired configurations. %a configuration of the form \eqref{eq:2 prop config}.

Start by choosing a propagator $p_1 = (j_1,j_1+2)$ of length 2 in $W$ (which exists by Remark~\ref{rem:props of length 2 and 3}).  If it is part of one of the configurations we are looking for then we are done, 
%a configuration \eqref{eq:2 prop config} we are done,
so suppose otherwise. 
By assumption: $p_1$ is not in configuration \ref{item pair of 2s}, there are no propagators of length $3$, and every vertex supports at least one propagator.  Therefore there must exist a propagator $q_1$ of length $\geq 4$ with one end on edge $j_1$ or on edge $j_1-1$ or on edge $j_1-2$.  Orient $q_1 = (i_1, k_1)$ such that  $p_1 \in \cP_{out}(q_1)$, (that is $(q_1, <_{i_1})$).

%This bounds a region $[i_1,k_1] = V_{out}(q_1) \cup V(q_1)$ on the outside of $(q_1, <_{i_1})$ (where $k_1 \in \{j_1-1, j_1, j_1+1\}$) which does not contain $p_1$.

@@

Call let $W_r := W|_{\cP_{in} \cup q_r}$. By the original hypotheses on $W$ every vertex in $v \in V_{in}(q_r)$ (which is a non-empty interval since $\ell(q_r) \geq 4$) must have support at least one propagator ($|\Prop(v)| \geq 1$), while the same is true for $V(q_r)$ because these vertices support $q_r$. By construction, the remaining vertices of $W|_{\cP_{in} \cup {q_r}}$, i.e. $V_{out}(q_r)$, are non-supporting.

By Remark~\ref{rem:props of length 2 and 3}, $W_r$ admits at least one propagator of length 2.  $W_r$ has no propagators of length $3$ since $W$ has none.  Let $p_{r+1}$ be a propagator of length 2 in $W_r$; if it forms part of configuration \ref{item big and 2} or \ref{item pair of 2s}
%the configuration \eqref{eq:2 prop config}
then we are done, so assume otherwise.

Note we may replace $q_r$ by any other propagator, $q'_{r}$ of length $\geq 4$ in $W_r$ such that $p_{r+1}$ and $q_r$ are on opposite sides; such a new $q'_{r}$ still satisfies all the hypotheses, and so, without loss of generality we may assume that $q_r$ has minimal length among propagators of length $\geq 4$ which have $p_{r+1}$ on their inside.

Write $p_{r+1} = (j_{r+1}, j_{r+1}+2)$.
There are two cases to consider.
The first case is that $i_r \leq j_{r+1} \leq i_r+2$ and $k_r-2 \leq j_{r+1}+2 \leq k_{r}$ (so $p_{r+1}$ has one end before $i_{r} + 3$ and the
other after $k_{r} - 2$).  Then since $p_{r+1}$ has length 2, it must be that $(i_r+3)+1 \geq k_r-2$ and so $q_r$ has length $\leq 6$.  By the minimality assumption on $q_r$ no propagator in $W_r$ has $q_r$ on one side and $p_{r+1}$ on the other side. Therefore, any propagator in $W_r$ other than $p_{r+1}$ itself must be of the form $(i,j)$ with $i_r\leq i, j\leq i_r+2$ or $k_r-2 \leq i,j\leq k_r$. That is, this propagator must be of length 2, and so we have configuration \ref{item pair of 2s} which we have already assumed does not occur.  Consequently, $\cP_{in}(q_r)$ contains only $p_{r+1}$ and so we have congifuratin  \ref{item big and 2} from the lemma statement which again we have assumed does not occur.
Therefore this first case cannot occur.

The second case is that one of the following three things happen (we continue to write $p_{r+1} = (j_{r+1}, j_{r+1}+2)$) 
\begin{itemize}
\item $i_r+3 \leq j_{r+1}\leq k_r-3$ or $i_r+3 \leq j_{r+1}+2\leq k_r-3$,
\item $i_r = j_{r+1}$, or 
\item $k_r-2 = j_{r+1}$
\end{itemize}
(Phrased more causually, this is that $p_{r+1}$ has either at least one end on an edge bounded by the vertices in the interval $[i_r+3, k_r-2]$ or both ends lie in $[i_r, i_r+3]$ or both ends lie in $[k_r-2, k_r+1]$.)
These situations all behave similarly.
By symmetry it suffices to only consider the second situation and the second possiblity of the first situation. Note that since $j_{r+1}+4 \leq k_r-1$ in both situations, and $j_{r+1}+4 \in V_{in}(q_r)$, thus $j_{r+1}+4$ is supported by a propagator in $W_r$ other than $q_r$.  Let $t$ be this propagator.  Since $p_{r+1}$ is not part of configuration \ref{item pair of 2s} and $W_r$ has no propagators of length 3, we must have that the length of $t$ is $\geq 4$.  If $q_r$ and $p_{r+1}$ were on different sides of $t$ then this would contradict the minimal length hypothesis on $q_r$.  Therefore $t$ has all the previous $q_i$ and $p_i$ on the same side and length $\geq 4$ and so we may set $q_{r+1} = t$ to continue the induction.

@@

The fact that Algorithm~\ref{alg:put GN on WLD} does construct the required Grassmann necklace is proved in Theorem~\ref{res:alg gives GN}. An worked example of Algorithm~\ref{alg:put GN on WLD} is given after Definition~\ref{def I_i as a function}.

\begin{algorithm}\label{alg:put GN on WLD}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram. This gives an algorithm for calculating the set $I_i$, for $i \in [n]$.

\begin{enumerate}
\item Fix a vertex $i \in [n]$. Set $a:=i$ and $I_i = \emptyset$.
\item While $\cP \neq \emptyset$, perform the following steps.
\begin{enumerate}
\item \textbf{Step $a$ for vertex $i$}: If $\Prop(a) \neq \emptyset$ in $W$, write $I_i = I_i \cup \{a\}$. Let $p \in \Prop(a)$ be the clockwise most propagator supported on $a$. Write $W = (\cP \setminus \{p\}, [n])$.
\item If $Prop(i) = \emptyset$ do nothing.
\item Increment $a$ by 1 and repeat from (a).
\end{enumerate}
\end{enumerate}
\end{algorithm}

If the algorithm assigns vertex $j$ to propagator $p$ from starting vertex $i$, we say that $p$ \emph{contributes} $j$ to $I_i$. Notationally, we represent this by allowing the $I_i$ symbol to represent a function as well as a set, as follows:
\begin{dfn}\label{def I_i as a function}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram. For each $i \in [n]$, define a function $I_i : \cP \longrightarrow [n]$ by
\[I_i(p) := \text{the vertex label that $p$ contributes to $I_i$ in Algorithm~\ref{alg:put GN on WLD},}\]
for each $p \in \cP$.
\end{dfn}

\begin{eg}Consider the admissible Wilson loop diagram
\[W\ =\ \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawlabeledprop{1}{0}{4}{0}{\footnotesize $q\ $}
	\drawlabeledprop{2}{0}{4}{-1}{\footnotesize $p$}
    \drawlabeledprop{5}{0}{8}{0}{\footnotesize $r\ \ $}
    \drawlabeledprop{5}{1}{7}{0}{\footnotesize $\ \ \ s$}

@@

To construct $I_1$ we start at vertex 1, so we begin with $i=1$, $a = 1$, and $I_1 = \emptyset$. 
\begin{itemize}
\item $a = 1$: Since $\Prop(1) = \{q,r\}$ and $r$ is the clockwise-most of these propagators, we let $I_1 = \{1\}$ and remove propagator $r$ from the diagram. 
\item $a = 2$: $\Prop(2) = \{p,q\}$, with $q$ being the clockwise-most of these propagators. So $I_1 = \{1,2\}$ and propagator $q$ is removed.
\item $a = 3$: $\Prop(3) = \{p\}$, so now $I_1 = \{1,2,3\}$ and propagator $p$ is removed. The only remaining propagator in the diagram is $s$.
\item $a = 4$: Since $p$ and $q$ were removed in earlier steps, we now have $\Prop(4) = \emptyset$.
\item $a = 5$: $\Prop(5) = \{s\}$, and we have $I_1 = \{1,2,3,5\}$.
\end{itemize}
There are no propagators left in the diagram, so the algorithm terminates and we have ${I_1 = \{1,2,3,5\}}$, or $1235$ for short. Viewed as a function, we have
\[I_1: \cP \longrightarrow [n] : \quad I_1(p) = 3,\quad I_1(q) = 2, \quad I_1(r) = 1, \quad I_1(s) = 5.\]
Applying the algorithm for all 8 starting vertices, we obtain the sets
\[1235, 2356, 3456, 4567, 5671, 6712, 7812, 8123.\]
The reader can easily verify that this sequence of $k$-sets satisfies the definition of a Grassmann necklace, and can (with significantly more work) also verify that this Grassmann necklace defines the positroid associated to the Wilson loop diagram $W$ given above.
\end{eg}

Given a propagator $p$ and a vertex $i$ in its support, it will be very useful in the following to understand on what set of vertices the Grassmann necklace algorithm assigns $i$ to $p$.  The answer is that the set is a non-empty cyclic interval.  Lemma~\ref{vertex cyclic int lem} establishes this, but first we need a preliminary lemma which is also useful in its own right.

@@

Since $I_i(p) \neq b$, there must exist a propagator $q$ inside of $p$ with $I_i(q) = b$. The propagator $q$ cannot end on the edge $(b-1,b)$, as this would contradict the minimality of $p$, so $q = (c,c+1,b,b+1)$ with $a <_i c <_i b$, and $I_i(q) = b$.

@@

Given a propagator $p$ of an admissible WLD $W$ and a vertex $i$ in the support of $p$, define
\[J_p^{(W)}(i) = \{m \in [n] \ | \ I^{(W)}_m(p) = i \},\]
i.e. the set of indices $m$ for which Algorithm~\ref{alg:put GN on WLD} assigns the value $i$ to the propagator $p$ in $W$. The following lemma establishes that these sets behave in a simple and predictable manner, a fact which we will repeatedly use in subsequent proofs.

@@

Let $p=(i,j)$ be a propagator of an admissible WLD $W$ on $n$ vertices.  Then $J_p^{(W)}(i)$, $J_p^{(W)}(i+1)$, $J_p^{(W)}(j)$, and $J_p^{(W)}(j+1)$ are each non-empty cyclic intervals which partition $[n]$ and occur in the given cyclic order.
\end{lem}

\begin{proof}
We will prove the result by induction on the number of propagators.  If $W$ has one propagator then the result is immediate.  Now suppose $W$ has more than one propagator.  Since non-supporting vertices have no effect on the Grassmann necklace algorithm, it suffices to prove the result for $W$ with $F(\emptyset) = \emptyset$.  Then by Lemma~\ref{lem sian}, $W$ has at least one of the four situations illustrated in Figure~\ref{fig 3 cases}.

@@

In each of the four cases, when we remove the propagator labelled $p$ we obtain a diagram which satisfies the statement of the theorem by the induction hypothesis, and contains a propagator ${q = (i,j)}$ with no other propagators inside it (although this region may or may not contain non-supporting vertices, which we will call $l, l+1, \ldots$ as necessary). Let $V$ be the diagram obtained by removing the propagator $p$ from $W$ (Figure \ref{fig V diagram}).

@@

\begin{thm}\label{res:alg gives GN}
The sequence of $k$-subsets $(I_1,\dots,I_n)$ obtained by applying Algorithm~\ref{alg:put GN on WLD} to all vertices of an admissible diagram $W$ is exactly the Grassmann necklace of $M(W)$.
\end{thm}

\begin{proof}
For each $i \in [n]$, let $I_i$ be the set of vertices assigned to the propagators of $W$ by Algorithm~\ref{alg:put GN on WLD} with starting vertex $i$. By Lemma~\ref{GN alg well defined}, we know that $|I_i| = k$ for each $i \in [n]$. It is easy to verify that the sequence $(I_1, \dots, I_n)$ is a Grassmann necklace if and only if $I_{i+1} \supseteq I_i \backslash \{i\}$ for all $i \in [n]$; this is the characterization we will use in this proof.

Suppose for a contradition that there exists an admissible diagram for which there exists an $i$ with $k\in I_i\setminus \{i\}$ and $k \not\in I_{i+1}$.  Fix $n$.  Let the triple $(W, i, k)$ be such a counterexample on $n$ vertices which is minimal with respect to the number of propagators. %, and among all those counterexamples on $n$ vertices with the same number of propagators, let $(W,i,k)$ be minimal with respect to the distance from $i$ to $k$ in the $\leq _i$ order.

If $i \not\in I_i$, then there are no propagators supported on $i$ at all.  In this case it is clear that applying Algorithm~\ref{alg:put GN on WLD} at vertex $i$ and vertex $i+1$ produces exactly the same result, i.e. $I_{i+1} = I_i$, and so $(W, i, k)$ is not a counterexample at all.

@@

\textbf{Case I}:  Suppose $p$ has one end on edge $i-1$.  Then $p$ is not supported on $i+1$, so in building $I_{i+1}$ we will take the same propagators as in the construction of $I_i$ from vertices $i+1$ up to $b-1$, that is $I_{i+1} \cap [i+1,b-1] = I_{i} \cap [i+1,b-1]$.  Furthermore, by Lemma~\ref{vertex cyclic int lem}, when building $I_{i+1}$ it must happen that $p$ is taken at vertex $b$, as otherwise $b$ would never be contributed by $p$.    Consequently, in building $I_{i+1}$, when the algorithm reaches vertex $b$ there cannot be any unassigned propagators remaining that are before $p$.  This is also true in when the algorithm constructing $I_i$ reaches $b$, since the same propagators have been assigned beforehand.  Finally, we also note that $k\geq_i b+1$.

Let $W'$ be the diagram obtained from $W$ by removing both $p$ and all propagators inside of $p$ (recall Definition~\ref{props inside p}).

By the above observations, if we commence Algorithm~\ref{alg:put GN on WLD} in $W'$ from vertex $b$, then we are in the same situation with respect to unassigned propagators as if we began at $i$ in $W$ and proceeded to $b$ following the algorithm; the propagators we assigned in the latter case are exactly the ones removed to build $W'$.  Similarly, starting at $i+1$ in $W$ and moving to $b+1$ leaves us in the same situation with respect to unassigned propagators as beginning at $b+1$ in $W'$ would.  This gives the equations
\begin{align*}
  I_i^{(W)} \cap [b,i-1] & = I_b^{(W')} \\
  I_{i+1}^{(W)} \cap [b+1,i-1] & = I_{b+1}^{(W')}
\end{align*}
where the diagram is indicated in the superscript.
Thus we have $k\in I_b^{(W')}\setminus\{b\}$ and $k\not\in I_{b+1}^{(W')}$, contradicting the minimality of $(W, i, k)$.

\textbf{Case II}: Suppose $p$ has one end on edge $i$.  Note that by assumption we have $I_i(p) = i$; this means that $p$ must be the first propagator lying on edge $i$, and hence we must have $I_{i+1}(p) = i+1$ as well.  Observe also that $k\geq_i i+2$ since $i+1\in I_{i+1}$.

Let $W'$ be the diagram obtained from $W$ by removing only the propagator $p$.  Then we have
\begin{align*}
  I_i^{(W)} \backslash \{i\} & = I_{i+1}^{(W')} \\
  I_{i+1}^{(W)} \backslash \{i+1\} & = I_{i+2}^{(W')}
\end{align*}
since in both cases the algorithm in $W'$ proceeds identically to that in $W$ after assigning $p$. Since $k \neq i+1$, we have $k\in I_{i+1}^{(W')}\setminus\{i+1\}$ but $k\not\in I_{i+2}^{(W')}$, contradicting the minimiality of $(W, i, k)$.

>>>>

We summarize here only the subset of positroid theory that we require in this paper; the interested reader is referred to \cite{Postnikov} for more details.

@@

Postnikov shows in \cite{Postnikov} that the positroids of $\Gr(k,n)$ are indexed by many different collections of objects, each with their own advantages and disadvantages. The two most suited to our purposes are Grassmann necklaces and Le diagrams, which we introduce below. In order to do so, we first need some preliminary definitions.

@@

In \cite{casestudy}, Agarwala and Fryer apply this process in the smallest non-trivial case: admissible Wilson loop diagrams with two propagators on six vertices. Even in this simple case, we see that the mapping from admissible WLD to positroid cells is neither one-to-one nor onto, and the process described above makes it almost impossible to track the relationship between the original Wilson loop diagram and the resulting positroid.

@@

%Given a propagator $p = (i, j)$ oriented $(p, <_i)$,  we can construct two obvious subdiagrams of $W$ by just removing all of the propagators on one side of $p$ and leaving the vertices intact, i.e. the subdiagrams
%\[  (\cP_{in}(p,<_i), [n])  \quad \textrm{and} \quad  (]cP_{out}(p,<_i), [n]).\]

We will inductively construct a sequence of pairs of propagators $(p_r,q_r)$ satisfying: $\ell(p_r) = 2$, and $p_r$ either forms part of configuration \ref{item big and 2} or \ref{item pair of 2s} from the statement,
  %a configuration \eqref{eq:2 prop config} in $D$
or there is a propagator $q_r$ satisfying
\begin{itemize}
\item $\ell(q_r) \geq 4$.
\item $\{p_1, \ldots, p_r\}$ and $\{q_1, \ldots, q_{r-1}\}$ are all on the same side of $q_r$.  
\end{itemize}
We can then choose the orientation of $q_r$ so that the previous $p_i$ and $q_i$ are all on the outside of $q_r$, and restrict our attention to the subdiagram $(\cP_{in}(q_r),[n])$. By the finiteness of $W$ this must eventually terminate in one of the desired configurations. %a configuration of the form \eqref{eq:2 prop config}.

Start by choosing a propagator $p_1 = (j_1,j_1+2)$ of length 2 in $W$ (which exists by Remark~\ref{rem:props of length 2 and 3}).  If it is part of one of the configurations we are looking for then we are done, 
%a configuration \eqref{eq:2 prop config} we are done,
so suppose otherwise.

From our existing assumptions, we know the following facts: $p_1$ is not in configuration \ref{item pair of 2s}, there are no propagators of length $3$, and every vertex supports at least one propagator.  Therefore there must be a propagator of length $\geq 4$ with one end on edge $j_1$ or on edge $j_1-1$ or on edge $j_1-2$.  Let $q_1 = (i_1, k_1)$ be this propagator, oriented such that $p_1 \in \cP_{out}(q_1,<_{i_1})$. This completes the base case.

@@

Let $W_r := (\cP_{in}(q_r),[n])$ be the subdiagram of $W$ consisting only of those propagators inside $q_r$ (including $q_r$ itself); in particular $W_r$ contains none of the previous $p_i$ or $q_i$. By the original hypotheses on $W$, every vertex in $[i_r,k_{r+1}]$ (which is a non-empty interval since $\ell(q_r) \geq 4$) must support at least one propagator. Therefore $W_r$ admits at least one propagator of length 2 (by Remark~\ref{rem:props of length 2 and 3}) and no propagators of length $3$ (by our assumption on $W$).

Let $p_{r+1} = (j_{r+1}, j_{r+1}+2)$ be a propagator of length 2 in $W_r$; if it forms part of configuration \ref{item big and 2} or \ref{item pair of 2s}
%the configuration \eqref{eq:2 prop config}
then we are done, so assume otherwise.

Note that we could replace $q_r$ by another propagator $q'_{r}$ of length $\geq 4$ in $W_r$, as long as $p_{r+1}$ and $q_r$ are on opposite sides of $q_r'$, and such a new $q'_{r}$ would still satisfy all of the induction hypotheses. Therefore without loss of generality we may assume that $q_r$ has minimal length among propagators of length $\geq 4$ in $W_r$ which have $p_{r+1}$ on their inside.

There are two cases to consider.

The first case is that $i_r \leq j_{r+1} \leq i_r+2$ and $k_r-2 \leq j_{r+1}+2 \leq k_{r}$ (so $p_{r+1}$ has one end before $i_{r} + 3$ and the other after $k_{r} - 2$, i.e. both ends of $p_{r+1}$ are close to the ends of $q_r$).  Then since $p_{r+1}$ has length 2, it must be that $(i_r+3)+1 \geq k_r-2$ and so $q_r$ has length $\leq 6$.  By the minimality assumption on $q_r$, no propagator in $W_r$ has $q_r$ on one side and $p_{r+1}$ on the other side. Therefore, any propagator in $W_r$ other than $q_r$ and $p_{r+1}$ must be of the form $(i,j)$ with $i_r\leq i < j\leq i_r+2$ or $k_r-2 \leq i < j\leq k_r$. Both of these require that $(i,j)$ has length 2, and so we have configuration \ref{item pair of 2s} which we have already assumed does not occur.  Consequently, $W_r$ contains only the two propagators $q_r$ and $p_{r+1}$, yielding configuration \ref{item big and 2} from the lemma statement and again contradicting our assumption. Therefore this first case cannot occur.

The second case is that (at least) one of the following four things happen:
\begin{itemize}
\item $i_r+3 \leq j_{r+1}\leq k_r-3$, or
\item $i_r+3 \leq j_{r+1}+2\leq k_r-3$, or
\item $i_r = j_{r+1}$, or 
\item $k_r-2 = j_{r+1}$
\end{itemize}
In other words, at least one end of $p_{r+1}$ is supported entirely by vertices in the interval $[i_r+3,k_r+2]$ (the first two bullet points) or both ends lie in $[i_r, i_r+3]$ (third point) or both ends lie in $[k_r-2, k_r+1]$ (fourth point). These situations all behave similarly, and by symmetry it suffices to only consider the second and third possibilities.

Note that since $j_{r+1}+4 \leq k_r-1$ in both of these situations, the vertex $j_{r+1}+4$ is inside $q_r$ but does not support either $p_r$ or $q_{r+1}$, and must therefore support some other propagator $t$.  Since $p_{r+1}$ is not part of configuration \ref{item pair of 2s} and $W_r$ has no propagators of length 3, it follows that $\ell(t) \geq 4$.  If $q_r$ and $p_{r+1}$ were on different sides of $t$ then this would contradict the minimal length hypothesis on $q_r$.  Therefore we can orient $t$ so that all previous $q_i$ and $p_i$ belong to $\cP_{out}(t)$, and set $q_{r+1} = t$ to continue the induction.

@@

The fact that Algorithm~\ref{alg:put GN on WLD} does construct the required Grassmann necklace is proved in Theorem~\ref{res:alg gives GN}. An worked example of Algorithm~\ref{alg:put GN on WLD} is given in Example~\ref{eg:apply GN alg}.

\begin{algorithm}\label{alg:put GN on WLD}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram. This gives an algorithm for constructing the set $I_j$, for $j \in [n]$.

\begin{enumerate}
\item Fix a vertex $j \in [n]$. Set $j:=i$ and $I_i = \emptyset$.
\item While $\cP \neq \emptyset$, perform the following steps:
\begin{enumerate}
\item \textbf{Step $j$ for vertex $i$}: If $\Prop(j) = \emptyset$, do nothing. Else, let $p \in \Prop(j)$ be the clockwise most propagator supported on $j$. Write $I_i = I_i\cup \{j\}$, and delete propagator $p$ from the diagram.
\item Increment $j$ by 1 and repeat from (a).
\end{enumerate}
\end{enumerate}
\end{algorithm}

Informally: one starts at $i$ and moves counterclockwise around the vertices of $W$, at each step removing the clockwise-most propagator supported on that vertex (if it exists). The set $I_i$ lists the vertices at which a propagator was deleted.

If the algorithm assigns vertex $j$ to propagator $p$ from starting vertex $i$, we say that $p$ \emph{contributes} $j$ to $I_i$. Notationally, we represent this by allowing the $I_i$ symbol to represent a function as well as a set, as follows:
\begin{dfn}\label{def I_i as a function}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram. For each $i \in [n]$, define a function $I_i^W : \cP \longrightarrow [n]$ by
\[I_i^W(p) := \text{the vertex label that $p$ contributes to $I_i$ in Algorithm~\ref{alg:put GN on WLD},}\]
for each $p \in \cP$. When the diagram $W$ is clear from context, we drop the superscript and simply write $I_i$.
\end{dfn}

\begin{eg}\label{eg:apply GN alg}Consider the admissible Wilson loop diagram
\[W\ =\ \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawlabeledprop{1}{0}{4}{0}{\footnotesize $q\ $}
	\drawlabeledprop{2}{0}{4}{-1}{\footnotesize $p$}
    \drawlabeledprop{5}{0}{8}{0}{\footnotesize $r\ \ $}
    \drawlabeledprop{5}{1}{7}{0}{\footnotesize $\ \ \ s$}

@@

To construct $I_1$ we start at vertex 1, so we begin with $i=1$, $m = 1$, and $I_1 = \emptyset$. 
\begin{itemize}
\item $j = 1$: Since $\Prop(1) = \{q,r\}$ and $r$ is the clockwise-most of these propagators, we let $I_1 = \{1\}$ and remove propagator $r$ from the diagram. 
\item $j = 2$: $\Prop(2) = \{p,q\}$, with $q$ being the clockwise-most of these propagators. So $I_1 = \{1,2\}$ and propagator $q$ is removed.
\item $j = 3$: $\Prop(3) = \{p\}$, so now $I_1 = \{1,2,3\}$ and propagator $p$ is removed. The only remaining propagator in the diagram is $s$.
\item $j = 4$: Since $p$ and $q$ were removed in earlier steps, we now have $\Prop(4) = \emptyset$.
\item $j = 5$: $\Prop(5) = \{s\}$, and we have $I_1 = \{1,2,3,5\}$.
\end{itemize}
There are no propagators left in the diagram, so the algorithm terminates and we have ${I_1 = \{1,2,3,5\}}$, or $1235$ for short. Viewed as a function, we have
\[I_1: \cP \longrightarrow [n] : \quad I_1(p) = 3,\quad I_1(q) = 2, \quad I_1(r) = 1, \quad I_1(s) = 5.\]
Applying the algorithm for all 8 starting vertices, we obtain the sets
\[1235, 2356, 3456, 4567, 5671, 6712, 7812, 8123.\]
The reader can easily verify that this sequence of $k$-sets satisfies the definition of a Grassmann necklace, and can (with significantly more work) also verify that this Grassmann necklace defines the positroid $M(W)$ associated to the Wilson loop diagram $W$ above.
\end{eg}

We first verify that this algorithm always terminates in fewer than $n$ steps.

@@

Since $I_i(p) \neq b$, there must exist a propagator $q$ inside of $(p,<_i)$ with $I_i(q) = b$. The propagator $q$ cannot end on the edge $(b-1,b)$, as this would contradict the minimality of $p$, so we have $q = (c,c+1,b,b+1)$ with $a <_i c <_i b$, and $I_i(q) = b$.

@@

We now reverse our perspective and ask: given a propagator $p$ and a vertex $i$ in the support of $p$, for which set of starting vertices does the algorithm assign $p$ to $i$?

Specifically, for each $i$ in the support of $p$, define the set
\[J_p^{(W)}(i) = \{m \in [n] \ | \ I^{(W)}_m(p) = i \},\]
the set of starting vertices $m$ such that $I_m^{(W)}$ assigns $p$ to $i$.

The following lemma establishes that these sets behave in a simple and predictable manner, a fact which we will repeatedly use in subsequent proofs.

@@

Let $p=(i,j)$ be a propagator in an admissible WLD $W$ on $n$ vertices.  Then $J_p^{(W)}(i)$, $J_p^{(W)}(i+1)$, $J_p^{(W)}(j)$, and $J_p^{(W)}(j+1)$ are each non-empty cyclic intervals which partition $[n]$ and occur in the given cyclic order.
\end{lem}

\begin{proof}
We will prove the result by induction on the number of propagators.  If $W$ has one propagator then the result is immediate.  Now suppose $W$ has more than one propagator.  Since non-supporting vertices have no effect on Algorithm~\ref{alg:put GN on WLD}, it suffices to prove the result for $W$ with $F(\emptyset) = \emptyset$.  Then by Lemma~\ref{lem sian}, $W$ has at least one of the four situations illustrated in Figure~\ref{fig 3 cases}.

@@

In each of the four cases, when we remove the propagator labelled $p$ we obtain a diagram which satisfies the statement of the theorem by the induction hypothesis, and contains a propagator ${q = (i,j)}$ with no other propagators inside it (with respect to the $<_i$ orientation), although this region may or may not contain non-supporting vertices which we will call $l, l+1, \ldots$ as necessary. Let $V$ be the diagram obtained by removing the propagator $p$ from $W$ (Figure \ref{fig V diagram}).

@@

\begin{thm}\label{res:alg gives GN}
The sequence of $k$-subsets $(I_1,\dots,I_n)$ obtained by applying Algorithm~\ref{alg:put GN on WLD} to all vertices of an admissible diagram $W$ is exactly the Grassmann necklace of $M(W)$.
\end{thm}
\note{Sian replaced all occurences of $k$ as an index with other letters: someone please proofread this specifically to check that none were missed! (then remove this note)}
\begin{proof}
For each $i \in [n]$, let $I_i$ be the set of vertices assigned to the propagators of $W$ by Algorithm~\ref{alg:put GN on WLD} with starting vertex $i$. By Lemma~\ref{GN alg well defined}, we know that $|I_i| = k$ for each $i \in [n]$. We first prove that $I_{i+1} \supseteq I_i \backslash \{i\}$ for all $i \in [n]$; this is equivalent to the definition of Grassmann necklace given in Definition~\ref{def:grassmann necklace}.

Suppose for a contradition that there exists an admissible diagram for which there exists an $i$ with $m\in I_i\setminus \{i\}$ and $m \not\in I_{i+1}$.  Fix $n$.  Let the triple $(W, i, m)$ be such a counterexample on $n$ vertices which is minimal with respect to the number of propagators. %, and among all those counterexamples on $n$ vertices with the same number of propagators, let $(W,i,k)$ be minimal with respect to the distance from $i$ to $k$ in the $\leq _i$ order.

If $i \not\in I_i$, then there are no propagators supported on $i$ at all.  In this case it is clear that applying Algorithm~\ref{alg:put GN on WLD} at vertex $i$ and vertex $i+1$ produces exactly the same result, i.e. $I_{i+1} = I_i$, and so $(W, i, m)$ is not a counterexample at all.

@@

\textbf{Case I}:  Suppose $p$ has one end on edge $i-1$.  Then $p$ is not supported on $i+1$, so in building $I_{i+1}$ we will take the same propagators as in the construction of $I_i$ from vertices $i+1$ up to $b-1$, that is $I_{i+1} \cap [i+1,b-1] = I_{i} \cap [i+1,b-1]$.  Furthermore, by Lemma~\ref{vertex cyclic int lem}, when building $I_{i+1}$ it must happen that $p$ is taken at vertex $b$, as otherwise $b$ would never be contributed by $p$.    Consequently, in building $I_{i+1}$, when the algorithm reaches vertex $b$ there cannot be any unassigned propagators remaining that are before $p$.  This is also true in when the algorithm constructing $I_i$ reaches $b$, since the same propagators have been assigned beforehand.  Finally, we also note that $m\geq_i b+1$.

Let $W' = (\cP_{out}(p,<_i), [n])$, i.e. the diagram obtained from $W$ by removing both $p$ and all propagators inside of $p$.

By the above observations, if we commence Algorithm~\ref{alg:put GN on WLD} in $W'$ from vertex $b$, then we are in the same situation with respect to unassigned propagators as if we began at $i$ in $W$ and proceeded to $b$ following the algorithm; the propagators we assigned in the latter case are exactly the ones removed to build $W'$.  Similarly, starting at $i+1$ in $W$ and moving to $b+1$ leaves us in the same situation with respect to unassigned propagators as beginning at $b+1$ in $W'$ would.  This gives the equations
\begin{align*}
  I_i^{(W)} \cap [b,i-1] & = I_b^{(W')} \\
  I_{i+1}^{(W)} \cap [b+1,i-1] & = I_{b+1}^{(W')}
\end{align*}
where the diagram is indicated in the superscript.
Thus we have $m\in I_b^{(W')}\setminus\{b\}$ and $m\not\in I_{b+1}^{(W')}$, contradicting the minimality of $(W, i, m)$.

\textbf{Case II}: Suppose $p$ has one end on edge $i$.  Note that by assumption we have $I_i(p) = i$; this means that $p$ must be the first propagator lying on edge $i$, and hence we must have $I_{i+1}(p) = i+1$ as well.  Observe also that $m\geq_i i+2$ since $i+1\in I_{i+1}$.

Let $W'$ be the diagram obtained from $W$ by removing only the propagator $p$.  Then we have
\begin{align*}
  I_i^{(W)} \backslash \{i\} & = I_{i+1}^{(W')} \\
  I_{i+1}^{(W)} \backslash \{i+1\} & = I_{i+2}^{(W')}
\end{align*}
since in both cases the algorithm in $W'$ proceeds identically to that in $W$ after assigning $p$. Since $m \neq i+1$, we have $m\in I_{i+1}^{(W')}\setminus\{i+1\}$ but $m\not\in I_{i+2}^{(W')}$, contradicting the minimiality of $(W, i, m)$.
